---
title: "DSCI Project R"
author: "Daniil Lenkov"
date: "3/18/2022"
output: html_document
---
# Dataset 1

# Hypothesis 1
Let’ read data fils and make some preprocessing:
* create special column in df1 for the futher analysis
* Rename ‘Year’ column in df2
* Make Deaths values in df2 as numeric
```{r}
df1 <- read.csv(file = 'D:/Daniil/Documents/DSCI_R/Project/final_1.csv')
df2 <- read.csv(file = 'D:/Daniil/Documents/DSCI_R/Project/final_2.csv')

df1['g_ng'] = 0
df1[df1$type.of.incident == 'Gunshot','g_ng'] = 'g'
df1[df1$type.of.incident != 'Gunshot','g_ng'] = 'ng'
colnames(df2)[1] <- 'Year'
df2['Deaths'] <- as.numeric(gsub(',','',unlist(df2['Deaths'])))


library(ggplot2)
library(ggthemes)
library(ggExtra)
```
It’s crucial to notice that dataset represent the fact of incidence, just the statistic without inner connections and reasons of crimes between each other (we can’t analyse it as time series, for example, or find correlation between variables)

First, let’s check, what’s the frequency of crimes based on fraction of BA degree in census tract

```{r}
#Mode function
get_mode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

#Histogram BA
p1_1 <- ggplot(df1, aes(x = college_level_census_tract)) +
  geom_histogram(color = 'black',
                 fill = 'rosybrown2',
                 bins = 30) + 
  geom_vline(aes(xintercept = get_mode(college_level_census_tract),
                 col='mode'),size=1) +
  geom_vline(aes(xintercept = mean(college_level_census_tract),
                 col='mean'),size=1) +
  scale_color_manual(name = "Average",
                     values = c(mode = "red",
                                mean = 'blue'))+
  labs(x  = 'Fraction of BA in census tract',
       title = 'Distribution of BA in census tract')+
  theme_classic()

plot(p1_1) 
```

Also we can check, what’s the distribution based on household income

```{r}
p1_2 <- ggplot(df1, aes(x = household_income_census_tract)) +
  geom_histogram(color = 'black',
                 fill = 'rosybrown2',
                 bins = 30) + 
  geom_vline(aes(xintercept = get_mode(household_income_census_tract),
                 col='mode'),size=1) +
  geom_vline(aes(xintercept = mean(household_income_census_tract),
                 col='mean'),size=1) +
  scale_color_manual(name = "Average",
                     values = c(mode = "red",
                                mean = 'blue'))+
  labs(x  = 'Household income, $',
       title = 'Distribution of household income in census tract')+
  theme_classic()

plot(p1_2)
```

But can we somehow merge both of plots above? The answer is yes. Moreover, we have an oppportunity to plot geom_point() as well

```{r}
p1_3 <-ggplot(df1,aes(x = household_income_census_tract,
                      y = college_level_census_tract)) +
  geom_point() +
  theme_classic()

ggMarginal(p1_3,type = 'histogram')
```

Another plot to represent the same idea of concenctretion of variables is density plot

```{r}
p2 <- ggplot(df1,aes(x = household_income_census_tract,
               y = college_level_census_tract)) +
  geom_point(size = 1) +
  geom_density_2d_filled(alpha = 0.9,
                         contour_var = 'ndensity',
                         bins = 7) +
  scale_fill_brewer(palette = 'YlOrRd') + 
  geom_density2d(color ='black',
                 bins = 7,
                 contour_var = 'ndensity') +
  theme_classic() + 
  theme(panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5)) + 
  labs(x = 'Household income in census tract',
       y = '% of people with BA in census tract',
       title = 'Household income vs BA')

plot(p2)
```

# Hypothesis 2

The idead of the second hypothesis is to check what the typical portrait of victim based on age,sex and murder weapon

```{r} 

#### Typical portrait of victim#####

#scale break
bins_10 <- function(df){
  cut_width(df$age_of_affected_individual, 
            10)}

p3 <- ggplot(data = df1, aes(x = type.of.incident))+
  geom_bar(aes(fill = bins_10(df1)))

p3 + facet_wrap(~gender_of_affected_individual +g_ng,scale = 'free') +
  scale_fill_discrete(name = "Age") +
  labs(x = 'Type of incident')+
  theme_classic()
```

# Dataset 2

# Hypothesis 1

Since we are in WV, it would be great to check the data for each disease in our state. For this, let’s plot separate geom_line() and smooth them. (I removed heart disease and cancer from here because of huge diffrence in Y axis; it’s possible to make cut of any of the axes but it’s better to do not with ggplot2 )

```{r}
p_2_1 <- ggplot(data = df2[(df2$State == 'West Virginia')&
                    (df2$Cause.Name != 'Cancer')&
                    (df2$Cause.Name != 'Heart disease')&
                    (df2$Cause.Name != 'All causes'),]) +
  geom_line(aes(x = Year,
                y = Deaths)) +
  geom_smooth(aes(x = Year,
                  y = Deaths), size = 1) +
  facet_wrap(~ Cause.Name) +
  theme_classic()

plot(p_2_1)
```

# Hypothesis 2

And, finally, we can represent the column plot and distribution of disease by year.

```{r}
p_2_2 <- ggplot(data = df2[(df2$State == 'West Virginia')&
                    (df2$Cause.Name != 'All causes'),], 
aes(x = Year,y = Deaths,fill = Cause.Name)) +
  geom_col() +
  theme_classic() +
  scale_fill_discrete(name = "Cause of Death")

plot(p_2_2)
```












