---
title: "DSCI_Final_Project"
author: "Daniil Lenkov"
date: '2022-05-04'
output: html_document
---
Let's import all necessary libraries
```{r}
#import of all libraries

library(tidyverse)

library(ggplot2)
library(ggthemes)
library(ggExtra)
library(ggalluvial)

library(alluvial)
library(magrittr)

library(cluster)
library(factoextra)

library(modelr)
```


# General trend of production and consumption of gas for the Europe 

First, we need to read and clean the data  

```{r}
# Consumption of gas in Europe for the last 50 years

df_gas_cons <- read.csv(file = 'C:/Users/Даниил/Desktop/DS_Europe_energy/gas_cons.csv')

df_gas_cons %<>% filter(Region %in% c('Total Europe')) %>%
  t(.) %>%
  as.double(.) %>%
  replace(is.na(.), 30) %>%
  tail(.,-5)

# Production of gas in Europe for the last 50 years
df_gas_prod <- read.csv(file = 'C:/Users/Даниил/Desktop/DS_Europe_energy/gas_prod.csv')

df_gas_prod %<>% filter(Region %in% c('Total Europe')) %>%
  t(.) %>%
  as.double(.) %>%
  replace(is.na(.), 100)

df_gas_prod_cons <- tibble(x = df_gas_prod, y = df_gas_cons)
df_gas_prod_cons
```

Now we can draw scatter plot. 
I wanted to pay special attention to the right concentration of plot:
*consumption stays the same despite the production change*

```{r}
ggplot(df_gas_prod_cons, aes(x=x, 
                             y=y)) +
  geom_point() +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5)) + 
  labs(x = 'Production',
       y = 'Consumption',
       title = 'Total Europe Gas Production and Consumption')
```

Trying to approximate this trend (which is a bit naive) I applied polynominal 
regression model

```{r}
#polynominal model
mod1 <- df_gas_prod_cons %>% 
  lm(formula =y ~ poly(x, 10))
summary(mod1)

#predicted values
grid <-  df_gas_prod_cons %>% 
  data_grid(x) %>% 
  gather_predictions(mod1)
```

And finally we can plot this model at the same chart
Right side is more flat, and it says about 
**critical need of Europe to substitute gas supplies to save a consumption**

```{r}
ggplot(df_gas_prod_cons, aes(x, y)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5)) + 
  labs(x = 'Production',
       y = 'Consumption',
       title = 'Total Europe Gas Production and Consumption')
```

# Alluvium plot for import and production of gas by country in europe

It's the most comprehensive dataset about energy in Europe
```{r}
#read the simplified energy balance file
df <- read.csv(file = 'C:/Users/Даниил/Desktop/DS_Europe_energy/nrg_bal_s_linear.csv')

dim(df)
```

Let's sort the data according to the conditions below 

Time Period = 2020

C0000X0350-0370 = *Solid fossil fuels* \
G3000 = *Natural gas* \
O4000XBIO = *Oil and petroleum products (excluding biofuel portion)* \
N900H = *Nuclear heat* \
RA000 = *Renewables and biofuels* \

Germany, UK, France, Italy, Spain, Portages, Netherlands, Poland, Sweden,
Belgium, Ireland

```{r}
eu_2020 <-df %>% filter(TIME_PERIOD == '2020',
                        siec %in% c('C0000X0350-0370',
                                    'G3000',
                                    'O4000XBIO',
                                    'N900H',
                                    'RA000'),
                        geo %in% c('DE','UK','FR',
                                   'IT','ES','NL',
                                   'PL','SE','BE','IE')) %>%
  select(c('geo','nrg_bal','siec','OBS_VALUE'))

imp <- eu_2020 %>% filter(nrg_bal == 'IMP')
prod <- eu_2020 %>% filter(nrg_bal == 'PPRD') 

s <- rbind(imp,prod) %>%
  select(geo,nrg_bal,siec,OBS_VALUE) %>%
  replace(is.na(.), 0)

head(s)
```

Quick test for compabiltiy with Alluvium plot
```{r}
is_alluvia_form(s, axes = 1:4, silent = TRUE)
```

And here is the Alluvium plot
Major economics of EU import majority of energy resourses
```{r}
ggplot(s,
       aes(y = OBS_VALUE, axis1 = nrg_bal, axis2 = geo)) +
  geom_alluvium(aes(fill = siec), width = 1/12) +
  geom_stratum(width = 1/12, fill = "black", color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Total Energy", "EU Country"), expand = c(.05, .05)) +
  theme_classic() +
  labs(y = 'GWH',
       title = 'European Reliance on Russian Gas Imports and on Gas for Energy, 2020') 
```

# Alluvium plot for import and production of gas by country in europe
Again, for the beginning we need to import, clean and transform the data
```{r}

#Import of natural gas by partner (Russia in this case) 

#1 MIO_M3 = 10,277,778 KWH = 10.277778 GWT hours 
#1 GWH = 1,000,000 KWH

df_gas <- read.csv(file = 'C:/Users/Даниил/Desktop/DS_Europe_energy/nrg_ti_gas__custom.csv')
df_gas_rus_2020 <- df_gas %>% 
  filter(TIME_PERIOD == '2020',
         siec == 'G3000') %>%
  mutate(gas_russia_gwh = OBS_VALUE*10.277778) %>%
  select(c(geo,gas_russia_gwh)) 

#Consumption of natural gas
df_gas_cons <- read.csv(file = 'C:/Users/Даниил/Desktop/DS_Europe_energy/nrg_cb_gas__custom.csv')
df_gas_cons_2020 <- df_gas_cons%>% 
  filter(TIME_PERIOD == '2020') %>%
  mutate(total_gas_cons = OBS_VALUE*10.277778) %>%
  select(c(geo,total_gas_cons)) 

#Gross inlang consumption of all energy, GWH
gic <- df %>%filter(nrg_bal == 'GIC',
                    TIME_PERIOD =='2020',
                    unit == 'GWH',
                    siec != 'TOTAL') %>%
  select(geo,OBS_VALUE)

gic %<>% group_by(geo) %>%
  summarize(total_energy_cons = sum(OBS_VALUE))

#Unite df
df_gas_cluster <- left_join(gic,df_gas_cons_2020,by = 'geo') %>%
  left_join(.,df_gas_rus_2020,by = 'geo')

df_gas_cluster
```

GDP parameter will be used futher for plotting 
```{r}
#GDP size of European Countries 
gdp <- read.csv(file = 'C:/Users/Даниил/Desktop/DS_Europe_energy/gdp.csv')
gdp %<>% rename(geo = ABBR)
head(gdp)
```

Calculation of ratios
```{r}
df_gas_cluster['share_rus_gas'] <- df_gas_cluster['gas_russia_gwh']/df_gas_cluster['total_gas_cons'] %>%
  replace(is.na(.), 0)

df_gas_cluster['share_gas_total'] <- df_gas_cluster['total_gas_cons']/df_gas_cluster['total_energy_cons'] %>%
  replace(is.na(.), 0)

df_gas_cluster  %<>%
  na.omit(share_rus_gas, share_gas_total) %>%
  filter(geo != 'EU27_2020',
         geo != 'EA19') %>%
  left_join(gdp,by = 'geo') %>%
  na.omit()

head(df_gas_cluster)
```

Creating separate df for clustering
```{r}
#Selected columns for clustering
df_gas_cluster1 <- df_gas_cluster %>%
  select(share_rus_gas, share_gas_total)

k2 <- kmeans(df_gas_cluster1, centers = 3)
k2
```

And finally we can plot our clusters
Here we can see that such economies as GE,IT,NL have great need in Russian 
gas import
```{r}
GDP_size = df_gas_cluster$GDP
Cluster = factor(k2$cluster)

ggplot(df_gas_cluster, aes(x=share_rus_gas, 
                                       y=share_gas_total, 
                                       color = Cluster,
                                       label = geo,
                                       size = GDP_size)) +
  geom_point() +
  geom_label() +
  theme_classic() + 
  geom_vline(aes(xintercept = 0.4),size=0.5) +
  geom_hline(aes(yintercept = 0.264),size=0.5) +
  theme(panel.border = element_rect(colour = "black", 
                                    fill=NA, 
                                    size=0.5)) + 
  labs(x = 'Share of Russian gus in gas consumption',
       y = 'Share of gas consumtion in total energy consumption',
       title = 'European Reliance on Russian Gas Imports and on Gas for Energy, 2020')

```
